---
title: "Predictors of Palliative Care in Colorectal Cancer Patients: A National Inpatient Sample Analysis"
subtitle: "Analysis for RCOP NIS-Apr-IM12"
author:
  - name: Eliza Aisha
    email: elizaaisha1@gmail.com
    affiliations: 
      - id: DMC
        name: Dow University of Health Sciences
        department: Department of Medicine
---


```{r}
#| label: libraries_setup
#| include: false

# Load libraries
library(tidyverse)
library(arrow)
library(survey)
library(gtsummary)
library(cardx)
library(broom.helpers)

# Load external dependencies
source("./scripts/labels.R")

```

```{r}
#| label: data_setup
#| include: false
#| cache: false

# Global options
TESTING <- F

#Import data
if (TESTING) {
  dta <- arrow::open_dataset("./data/NIS_Palliative_CRC_testing.parquet") |> collect()
} else {
  dta <- arrow::open_dataset("./data/NIS_Palliative_CRC.parquet/") |> collect()
}

# Set factor references
dta <- dta |>
  mutate(
    Insurance = fct_relevel(Insurance, "Private"),
    RACE = fct_relevel(RACE, "White"),
    Palliative_Care = fct_relevel(Palliative_Care, "No")
)

# Set survey options for lonely PSUs
options(survey.lonely.psu = "adjust")
options(survey.adjust.domain.lonely = TRUE)

# Define survey design
dsgn <- svydesign(
  id = ~ HOSPID,
  strata = ~ NIS_STRATUM,
  weights = ~ TRENDWT,
  data = dta,
  nest = TRUE
)

# Subset data for specific condition
dsgn <- subset(dsgn, Colorectal_CA == "Yes" & !is.na(Palliative_Care))

# Clean up memory
rm(dta)
gc()


```

## Preamble:

-   **Reference Papers:**

    -   [Mojtahedi et al. 2021](https://pmc.ncbi.nlm.nih.gov/articles/PMC8496534/#s0002)

-   **Study Objective**:  To characterize the demographic, clinical, and hospital-level factors associated with utilization of palliative care among patients with colorectal cancer.

-   **Data Source**: Cross-sectional analysis of the National Inpatient Sample (NIS) from 2018 to 2020. The NIS is part of the Healthcare Cost and Utilization Project (HCUP) and provides a stratified sample of inpatient discharges from U.S. community hospitals across all payers.

-   **Cohort Definition:** Patients were included if they met the following criteria

    -   Age ≥18 years

    -   Diagnosis of colorectal cancer

    -   Non-missing data for palliative care consultation status
    
-   **Outcome of Interest:** Receipt of palliative care consultation, defined by ICD-10-CM code Z51.5 ("Encounter for palliative care") in any diagnosis field.

-   **Covariates for Adjustment:** All models adjusted for

    -   **Demographic Factors:** Age, sex, race/ethnicity

    -   **Socioeconomic Factors:** Primary payer, ZIP code–level income quartile

    -   **Clinical Factors:** Elixhauser comorbidity index (aggregate score), APR-DRG risk of mortality

    -   **Hospital-Level Factors:** Hospital region, bed size, teaching status, and urban/rural location
    
-   **Statistical Methods:**

    -   **Descriptive Analysis:**

        -   Patient and hospital characteristics were compared between those who did and did not receive palliative care consultation.

        -   Categorical variables: Compared using Pearson’s chi-square test with Rao & Scott adjustment

        -   Continuous variables: Analyzed using the design-based Kruskal–Wallis test, a non-parametric alternative suitable for survey-weighted data

    -   **Multivariable Analysis:**

        -   Predictors of palliative care consultation were assessed using survey-weighted logistic regression with the `quasibinomial` family.

        -   Results were reported as adjusted odds ratios (aORs) with 95% confidence intervals.

-   **Software:** All statistical analyses were performed using the R Statistical Language (Version 4.5.0; R Foundation for Statistical Computing, Vienna, Austria), incorporating survey-weighted procedures via the *survey* package to account for the complex sampling design of NIS.

## Baseline Table:

```{r}
#| label: baseline
#| cache: false

# Create baseline table
tbl_svybaseline <- dsgn %>%
  tbl_svysummary(
    by = Palliative_Care,
    include = baseline_var,
    label = baseline_var_labels,
    missing = "no"
  ) %>%
  add_p() |>
  add_overall()

# Print baseline table
tbl_svybaseline

```

## Multivariable logistic regression:

### Palliative care consultation:

```{r}
#| label: reg_Palliative_Care

# Palliative_Care
#Create the regression model
reg_model <- svyglm(
  Palliative_Care ~
    AGE +
      FEMALE +
      RACE +
      Insurance +
      ZIPINC_QRTL +
      Hosp_Census_Region +
      HOSP_BEDSIZE +
      HOSP_LOCTEACH +
      APRDRG_Risk_Mortality +
      elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model, exponentiate = T, label = reg_var_labels)

```
