---
title: "Palliative Care in Colorectal Cancer Patients: A National Inpatient Sample Analysis"
subtitle: "Analysis for RCOP NIS-Apr-IM12"
author:
  - name: Eliza Aisha
    email: elizaaisha1@gmail.com
    affiliations: 
      - id: DMC
        name: Dow University of Health Sciences
        department: Department of Medicine
---


```{r}
#| label: libraries_setup
#| include: false

# Load libraries
library(tidyverse)
library(arrow)
library(survey)
library(gtsummary)

# Load external dependencies
source("./scripts/labels.R")

```

```{r}
#| label: data_setup
#| include: false
#| cache: false

# Global options
TESTING <- F

#Import data
if (TESTING) {
  dta <- arrow::open_dataset("./nis_data/NIS_Pall_CRC_testing.parquet") |> collect()
} else {
  dta <- arrow::open_dataset("./nis_data/NIS_Pall_CRC.parquet/") |> collect()
}

# Set factor references
dta <- dta |>
  mutate(
    IE_Afib = fct_relevel(IE_Afib, "Infective Endocarditis without Atrial Fibrillation"),
    Insurance = fct_relevel(Insurance, "Private"),
    RACE = fct_relevel(RACE, "White"),
    DIED = as.factor(DIED),
    LOS = as.numeric(LOS),
    ARF = as.factor(ARF),
    AIS = as.factor(AIS),
    AHF = as.factor(AHF)
  )

# Set survey options for lonely PSUs
options(survey.lonely.psu = "adjust")
options(survey.adjust.domain.lonely = TRUE)

# Define survey design
dsgn <- svydesign(
  id = ~ HOSPID,
  strata = ~ NIS_STRATUM,
  weights = ~ TRENDWT,
  data = dta,
  nest = TRUE
)

# Subset data for specific condition
dsgn <- subset(dsgn, IE == "Yes")

# Clean up memory
rm(dta)
gc()


```

## Preamble:

-   **Reference Papers:**

    -   [Sidhu et al. 2022](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9207776/pdf/main.pdf)

-   **Study Objective**: Examine the utilization of palliative care among patients with colorectal cancer.

-   **Data Source**: Cross-sectional analysis of the National Inpatient Sample (NIS) from 2018 to 2020.

-   **Patient Selection**: Included all inpatient admissions with a diagnosis of colorectal cancer.
    
-   **Statistical Analysis**: Multiple logistic and linear regression to determine the independent association of atrial fibrillation with in-hospital outcomes, respectively, adjusted for:

    -   Demographics: Age, sex, race, residential income, insurance, and hospital region,

    -   Medical History: Hypertension, diabetes mellitus, hyperlipidemia, coronary artery disease, peripheral vascular disease, heart failure, chronic kidney disease, chronic obstructive pulmonary disease, liver disease, obstructive sleep apnea, anemia, obesity, alochol use disorder, and tobacco use disorder.

    -   Comorbidities: Charlson comorbidity index.

-   **Software:** All analyses were performed using R Version 4.4.1 (R Foundation for Statistical Computing, Vienna, Austria)

## Baseline Table:

```{r}
#| label: baseline
#| cache: true

# Create baseline table
tbl_svybaseline <- dsgn %>%
  tbl_svysummary(
    by = IE_Afib,
    include = baseline_var,
    label = baseline_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p() %>%
  add_overall()

# Print baseline table
tbl_svybaseline

```

## Outcomes Table:

```{r}
#| label: outcomes_table

# Create outcomes table
tbl_svyoutcomes <- dsgn %>%
  tbl_svysummary(
    include = outcome_var,
    label = outcome_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p() %>%
  add_overall()

# Print outcomes table
tbl_svyoutcomes

```

## Multivariable Logistic Regression:

### Palliative Care Utilization:

```{r}
#| label: reg_pall_care

# Create the regression model for palliative care
reg_model_pall_care <- svyglm(
  Pall_Care ~ AGE + FEMALE + RACE + ZIPINC_QRTL + APRDRG_Severity + Insurance + charlindex + Hosp_Census_Region,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_pall_care, exponentiate = T, label = reg_var_labels)

```

### Discharge Disposition:

```{r}
#| label: ARF

# Create the regression model for ARF
reg_model_Disp <- svyglm(
  DISPUNIFORM ~ Pall_care + AGE + FEMALE + RACE + ZIPINC_QRTL + APRDRG_Severity + Insurance + charlindex + Hosp_Census_Region,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_Disp, exponentiate = T, label = reg_var_labels)

```

