---
title: "The Association of Malignancy With Vascular Outcomes in Atrial Fibrillation (Sensitivity Analysis excluding Patients with Prior Stroke)"
subtitle: "Analysis for RCOP 2025-NIS-13-Internal Medicine"
author:
  - name: Eliza Aisha
    email: elizaaisha1@gmail.com
    affiliations: 
      - id: DMC
        name: Dow University of Health Sciences
        department: Department of Medicine
---


```{r}
#| label: libraries_setup
#| include: false

# Load libraries
library(tidyverse)
library(arrow)
library(survey)
library(gtsummary)

# Load external dependencies
source("./scripts/labels.R")

```

```{r}
#| label: data_setup
#| include: false
#| cache: false

# Global options
TESTING <- F

#Import data
if (TESTING) {
  dta <- arrow::open_dataset("./data/NIS_CA_Afib_testing.parquet") |> collect()
} else {
  dta <- arrow::open_dataset("./data/NIS_CA_Afib.parquet") |> collect()
}

# Set factor references
dta <- dta |>
  mutate(
    CA_Afib = fct_relevel(CA_Afib, "Atrial Fibrillation without Cancer"),
    Insurance = fct_relevel(Insurance, "Private"),
    RACE = fct_relevel(RACE, "White"),
    DIED = as.factor(DIED),
    AMI = as.factor(AMI),
    AIS = as.factor(AIS),
    VTE = as.factor(VTE),
    MB = as.factor(MB),
    LOS = as.numeric(LOS)
  )

# Set survey options for lonely PSUs
options(survey.lonely.psu = "adjust")
options(survey.adjust.domain.lonely = TRUE)

# Define survey design
dsgn <- svydesign(
  id = ~ HOSPID,
  strata = ~ NIS_STRATUM,
  weights = ~ TRENDWT,
  data = dta,
  nest = TRUE
)

# Subset data for specific condition
dsgn <- subset(dsgn, A_Fib == "Yes" & pStroke == "No" & CatheterAblation == "No" & AGE > 18 & !is.na(CA_Afib))

# Clean up memory
rm(dta)
gc()


```

## Preamble:

-   **Reference Papers:**

    -   [Elbadawi et al. 2017](https://pubmed.ncbi.nlm.nih.gov/29352566/)
    -   [El-Rayes et al. 2025](https://www.jacc.org/doi/10.1016/j.jaccao.2024.10.014)

## Preamble:

-   **Study Objective**: To examine the impact of malignancy on outcomes among admissions with atrial fibrillation.

-   **Data Source**: Cross-sectional analysis of the National Inpatient Sample (NIS) from 2018 to 2020. The NIS is part of the Healthcare Cost and Utilization Project (HCUP) and provides a stratified sample of inpatient discharges from U.S. community hospitals across all payers.

-   **Cohort Definition:** Patients were included if they met the following criteria

    -   Age ≥18 years

    -   Had a diagnosis of atrial fibrillation and cancer
    
    -   Patients undergoing catheter ablation were excluded

-   **Primary Outcomes**:

    -   Acute ischemic stroke
    
    -   Venous thromboembolism
            
    -   Major bleeding
    
-   **Secondary Outcomes**:
    
    -   In-hospital all-cause mortality
    
    -   Total hospital length of stay (days)

    -   Total charge, inflation adjusted to 2020 (\$)
    
-   **Covariates for Adjustment:** All models adjusted for the following

    -   Demographics: Age, sex, race, income, insurance, location of patient's residence, hospital region, hospital bedsize, hospital location, and teaching status.

    -   Medical History: Hypertension, diabetes mellitus, obesity, hyperlipidemia, coronary artery disease, peripheral vascular disease, heart failure, dementia, chronic kidney disease, chronic pulmonary disease, obstructive sleep apnea, liver disease, valvular heart disease, valvular heart disease, alcohol abuse, smoker, prior cardiac surgery, and type of cancer.

    -   Comorbidities: Elixhauser comorbidity index.

-   **Statistical Methods:**

    -   **Univariable Analysis:**

        -   Categorical variables: Compared using Pearson’s chi-square test with Rao & Scott adjustment

        -   Continuous variables: Analyzed using the design-based Kruskal–Wallis test, a non-parametric alternative suitable for survey-weighted data

    -   **Multivariable Analysis:**

        -   Binary outcomes (e.g., mortality): Survey-weighted logistic regression using the `quasibinomial` family

        -   Continuous outcomes (e.g., LOS, hospital charges): Survey-weighted linear regression using the `gaussian` family

        -   Results reported as adjusted odds ratios (aORs) or beta coefficients with 95% confidence intervals

-   **Software:** All statistical analyses were performed using the R Statistical Language (Version 4.5.0; R Foundation for Statistical Computing, Vienna, Austria), incorporating survey-weighted procedures via the *survey* package to account for the complex sampling design of NIS.

## Baseline Table:

```{r}
#| label: baseline
#| cache: false

# Create baseline table
tbl_svybaseline <- dsgn %>%
  tbl_svysummary(
    by = CA_Afib,
    include = baseline_var,
    label = baseline_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p() %>%
  add_overall()

# Print baseline table
tbl_svybaseline

# Create outcomes table
tbl_svyoutcomes <- dsgn %>%
  tbl_svysummary(
    by = CA_Afib,
    include = outcome_var,
    label = outcome_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p()

# Print outcomes table
tbl_svyoutcomes

```

## Multivariable Logistic Regression:

### All-Cause Mortality:

```{r}
#| label: reg_all_cause_mortality

# Create the regression model for all cause mortality
reg_model_died <- svyglm(
  DIED ~ CA_Afib + AGE + FEMALE + RACE + elixsum + ZIPINC_QRTL + Insurance + Hosp_Census_Region + PL_NCHS + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + VHD + Liver_Dis + OSA + Anemia + Alcohol + Obesity + pCardiacSurg + Smoker,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_died, exponentiate = T, label = reg_var_labels)

```

### Acute Ischemic stroke:

```{r}
#| label: reg_all_AIS

# Create the regression model for AIS
reg_model_AIS <- svyglm(
  AIS ~ CA_Afib + AGE + FEMALE + RACE + elixsum + ZIPINC_QRTL + Insurance + Hosp_Census_Region + PL_NCHS + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + VHD + Liver_Dis + OSA + Anemia + Alcohol + Obesity + pCardiacSurg + Smoker,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_AIS, exponentiate = T, label = reg_var_labels)

```

### Venous Thromboembolism:

```{r}
#| label: reg_all_DVT

# Create the regression model for DVT
reg_model_VTE <- svyglm(
  VTE ~ CA_Afib + AGE + FEMALE + RACE + elixsum + ZIPINC_QRTL + Insurance + Hosp_Census_Region + PL_NCHS + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + VHD + Liver_Dis + OSA + Anemia + Alcohol + Obesity + pCardiacSurg + Smoker,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_VTE, exponentiate = T, label = reg_var_labels)

```

### Major bleeding:

```{r}
#| label: reg_all_MB

# Create the regression model for MB
reg_model_MB <- svyglm(
  MB ~ CA_Afib + AGE + FEMALE + RACE + elixsum + ZIPINC_QRTL + Insurance + Hosp_Census_Region + PL_NCHS + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + VHD + Liver_Dis + OSA + Anemia + Alcohol + Obesity + pCardiacSurg + Smoker,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_MB, exponentiate = T, label = reg_var_labels)

```

## Multivariable Linear Regression:

### Length of Stay:

```{r}
#| label: reg_length_of_stay

# Create the regression model
reg_model_LOS <- svyglm(
  LOS ~ CA_Afib + AGE + FEMALE + RACE + elixsum + ZIPINC_QRTL + Insurance + Hosp_Census_Region + PL_NCHS + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + VHD + Liver_Dis + OSA + Anemia + Alcohol + Obesity + pCardiacSurg + Smoker,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model_LOS, exponentiate = F, label = reg_var_labels)

```

### Inflation Adjusted Total Charge:

```{r}
#| label: reg_total_charge

# Create the regression model for Total Charge (Inflation Adjusted)
reg_model_TOTCHG <- svyglm(
  adj_TOTCHG ~ CA_Afib + AGE + FEMALE + RACE + elixsum + ZIPINC_QRTL + Insurance + Hosp_Census_Region + PL_NCHS + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + VHD + Liver_Dis + OSA + Anemia + Alcohol + Obesity + pCardiacSurg + Smoker,
  design = dsgn,
  family = "gaussian"
  )

# Print regression table
tbl_regression(reg_model_TOTCHG, exponentiate = F, label = reg_var_labels)

```
