---
title: "Impact of BMI on In-Hospital Outcomes in Colorectal Cancer Patients"
subtitle: "Analysis for RCOP NIS Gastro2"
author:
  - name: Eliza Aisha
    email: elizaaisha1@gmail.com
    affiliations: 
      - id: DMC
        name: Dow University of Health Sciences
        department: Department of Medicine
---


```{r}
#| label: libraries_setup
#| include: false

# Load libraries
library(tidyverse)
library(arrow)
library(survey)
library(gtsummary)
library(cardx)
library(broom.helpers)

# Load external dependencies
source("./scripts/labels.R")

```

```{r}
#| label: data_setup
#| include: false
#| cache: false

# Global options
TESTING <- F

# Import data
if (TESTING) {
  dta <- arrow::open_dataset("./nis_data/NIS_Gastro2_testing.parquet") |> collect()
} else {
  dta <- arrow::open_dataset("./nis_data/NIS_Gastro2.parquet/") |> collect()
}

# Set factor references
dta <- dta |>
  mutate(
    BMI = fct_relevel(BMI, BMI_order),
    DIED = fct_relevel(DIED, "No"),
    LOS = as.numeric(LOS),
    adj_TOTCHG = as.numeric(adj_TOTCHG)
  )

# Set survey options for lonely PSUs
options(survey.lonely.psu = "adjust")
options(survey.adjust.domain.lonely = TRUE)

# Define survey design
dsgn <- svydesign(
  id = ~HOSPID,
  strata = ~NIS_STRATUM,
  weights = ~TRENDWT,
  data = dta,
  nest = TRUE
)

# Subset data for specific condition
dsgn <- subset(dsgn, !is.na(Colo_Rect_CA) & !is.na(BMI))

# Clean up memory
rm(dta)
gc()

```

## Preamble:

-   **Reference Papers:**

    -   [Salman et al. 2024](https://journals.lww.com/coronary-artery/abstract/9900/obesity_paradox_in_coronary_artery_disease_.331.aspx)
    -   [Persaud et al.2019](https://www.ahajournals.org/doi/full/10.1161/STROKEAHA.119.024638)
    -   [Sreenivasan et al. 2022](https://doi.org/10.1016/j.amjcard.2022.04.011)

-   **Study Objective**: Examine the impact of body mass index (BMI) on outcomes among inpatient admissions with colorectal cancer.

-   **Data Source**: Cross-sectional analysis of the National Inpatient Sample (NIS) from 2018 to 2020.

-   **Patient Selection**: Included all inpatient admissions with colorectal cancer.

-   **BMI Categories**:

    -   Underweight: BMI ≤19.9 kg/m²

    -   Normal Weight: BMI 20.0 to 24.9 kg/m²

    -   Overweight: BMI 25.0 to 29.9 kg/m²

    -   Class I Obesity: BMI 30.0 to 34.9 kg/m²

    -   Class II Obesity: BMI 35.0 to 39.9 kg/m²

    -   Class III Obesity: BMI ≥40.0 kg/m²

-   **Outcomes**:

    -   In-hospital all-cause mortality

    -   Total hospital length of stay (days)

    -   Total charge, inflation adjusted to 2020 (\$)

-   **Statistical Analysis**: Multiple logistic and linear regression to determine the independent association of BMI categories with in-hospital mortality and length of stay, respectively, adjusted for:

    -   Demographics: Age, sex, race, residential income, and insurance.

    -   Medical History: Cardiac arrhythmia, Anemia, Cerebrovascular disease, Chronic kidney disease, Chronic obstructive pulmonary disease, Diabetes, Dementia, Heart failure, Hypertension, Hyperlipidemia, Myocardial infarction, Peripheral vascular disease, Rheumatoid disease, Valvular heart disease
    
    -   Comorbidities: Charlson comorbidity index.

-   **Software:** All analyses were performed using R Version 4.4.1 (R Foundation for Statistical Computing, Vienna, Austria)

## Baseline Table:

```{r}
#| label: baseline

# Create baseline table
tbl_svybaseline <- dsgn %>%
  tbl_svysummary(
    by = BMI,
    include = baseline_var,
    label = baseline_var_labels,
    missing = "no"
  ) %>%
  add_p() |>
  add_overall()

# Print baseline table
tbl_svybaseline

```

## Outcomes Table:

```{r}
#| label: outcomes_table

# Create outcomes table
tbl_svyoutcomes <- dsgn %>%
  tbl_svysummary(
    by = BMI,
    include = outcome_var,
    label = outcome_var_labels,
    missing = "no"
  ) %>%
  add_p()

# Print outcomes table
tbl_svyoutcomes

```

## Multivariable Logistic Regression:

### All-Cause Mortality:

```{r}
#| label: reg_all_cause_mortality

#Create the regression model
reg_model <- svyglm(
  DIED ~ BMI + AGE + FEMALE + RACE + ZIPINC_QRTL + Insurance + charlindex + Arrhythmia + Anemia + CEVD + CKD + COPD + DM + Dementia + HF + HTN + Hyperlip + MI + PVD + Rheumatoid + Valvular_Disease,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model, exponentiate = T, label = reg_var_labels)

```

## Multivariable Linear Regression:

### Length of Stay:

```{r}
#| label: reg_length_of_stay

# Create the regression model
reg_model <- svyglm(
  LOS ~ BMI + AGE + FEMALE + RACE + ZIPINC_QRTL + Insurance + charlindex + Arrhythmia + Anemia + CEVD + CKD + COPD + DM + Dementia + HF + HTN + Hyperlip + MI + PVD + Rheumatoid + Valvular_Disease,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model, exponentiate = F, label = reg_var_labels)


```

### Inflation Adjusted Total Charge:

```{r}
#| label: reg_total_charge

# Create the regression model
reg_model <- svyglm(
  adj_TOTCHG ~ BMI + AGE + FEMALE + RACE + ZIPINC_QRTL + Insurance + charlindex + Arrhythmia + Anemia + CEVD + CKD + COPD + DM + Dementia + HF + HTN + Hyperlip + MI + PVD + Rheumatoid + Valvular_Disease,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model, exponentiate = F, label = reg_var_labels)

```
