---
title: "Clinical Characteristics and Outcomes of Patients with Inflammatory Bowel Disease and Acute Myocardial Infarction"
author:
  - name: Eliza Aisha
    email: elizaaisha1@gmail.com
    affiliations: 
      - id: DMC
        name: Dow University of Health Sciences
        department: Department of Medicine
        
  - name: Mariam Shahabi
    affiliations: 
      - id: DMC
        name: Dow University of Health Sciences
        department: Department of Medicine      
---

```{r}
#| label: libraries_setup
#| include: false

# Load libraries
library(tidyverse)
library(arrow)
library(survey)
library(gtsummary)

# Load external dependencies
source("./scripts/labels.R")

```

```{r}
#| label: data_setup
#| include: false
#| cache: false

# Global options
TESTING <- F

#Import data
if (TESTING) {
  dta <- arrow::open_dataset("./nis_data/NIS_AMI_IBD_testing.parquet") |> collect()
} else {
  dta <- arrow::open_dataset("./nis_data/NIS_AMI_IBD.parquet/") |> collect()
}

# Set factor references
dta <- dta |>
  mutate(
    IBD = fct_relevel(IBD, "AMI without IBD"),
    Insurance = fct_relevel(Insurance, "Private"),
    RACE = fct_relevel(RACE, "White"),
    DIED = fct_relevel(DIED, "No"),
    LOS = as.numeric(LOS),
    adj_TOTCHG = as.numeric(adj_TOTCHG),
    CardioShock = as.factor(CardioShock),
    CardiacArrest = as.factor(CardiacArrest),
    Stroke = as.factor(Stroke),
    PCIProcedure = as.factor(PCIProcedure),
    CABGProcedure = as.factor(CABGProcedure)
  )

# Set survey options for lonely PSUs
options(survey.lonely.psu = "adjust")
options(survey.adjust.domain.lonely = TRUE)

# Define survey design
dsgn <- svydesign(
  id = ~ HOSPID,
  strata = ~ NIS_STRATUM,
  weights = ~ TRENDWT,
  data = dta,
  nest = TRUE
)

# Subset data for specific condition
dsgn <- subset(dsgn, AMI == "Yes" & AGE >= 18)

# Clean up memory
rm(dta)
gc()

```


## Preamble:

-   **Study Objective**: Examine the impact of irritable bowel disease on outcomes among inpatient admissions with acute myocardial infarction.

-   **Data Source**: Cross-sectional analysis of the National Inpatient Sample (NIS) from 2018 to 2020.

-   **Patient Selection**: Included all inpatient admissions with a principal diagnosis of acute myocardial infarction.

-   **Primary Outcomes**:

    -   In-hospital all-cause mortality

    -   Total hospital length of stay (days)

    -   Total charge, inflation adjusted to 2020 (\$)
    

-   **Secondary Outcomes**:

    -   Stroke

    -   Cardiogenic Shock

    -   Cardiac Arrest

    -   Utilization of Percutaneous Coronary Intervention

    -   Utilization of Coronary Artery Bypass Grafting
    

-   **Covariates for Adjustment:** All models adjusted for the following

    -   Demographics: Age, sex, race, income, insurance, urbanicity, hospital region, hospital bedsize, hospital location, and teaching status.

    -   Medical History: Hypertension, diabetes mellitus, obesity, hyperlipidemia, coronary artery disease, peripheral vascular disease, heart failure, dementia, cancer, chronic kidney disease, chronic pulmonary disease, obstructive sleep apnea, dementia, liver disease, valvular heart disease, prior stroke, prior cardiac surgery, presence of pacemaker/defibrillator.

    -   Comorbidities: Elixhauser comorbidity index.

-   **Statistical Methods:**

    -   **Univariable Analysis:**

        -   Categorical variables: Compared using Pearson’s chi-square test with Rao & Scott adjustment

        -   Continuous variables: Analyzed using the design-based Kruskal–Wallis test, a non-parametric alternative suitable for survey-weighted data

    -   **Multivariable Analysis:**

        -   Binary outcomes (e.g., mortality): Survey-weighted logistic regression using the `quasibinomial` family

        -   Continuous outcomes (e.g., LOS, hospital charges): Survey-weighted linear regression using the `gaussian` family

        -   Results reported as adjusted odds ratios (aORs) or beta coefficients with 95% confidence intervals

-   **Software:** All statistical analyses were performed using the R Statistical Language (Version 4.5.0; R Foundation for Statistical Computing, Vienna, Austria), incorporating survey-weighted procedures via the *survey* package to account for the complex sampling design of NIS.

## Baseline Table:

```{r}
#| label: baseline
#| cache: false

# Create baseline table
tbl_svybaseline <- dsgn %>%
  tbl_svysummary(
    by = IBD,
    include = baseline_var,
    label = baseline_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p() %>%
  add_overall()

# Print baseline table
tbl_svybaseline

```

## Outcomes Table:

```{r}
#| label: outcomes_table

# Create outcomes table
tbl_svyoutcomes <- dsgn %>%
  tbl_svysummary(
    by = IBD,
    include = outcome_var,
    label = outcome_var_labels,
    missing = "no"
  ) %>%
  add_p()

# Print outcomes table
tbl_svyoutcomes

```

## Multivariable Logistic Regression:

### All-Cause Mortality:

```{r}
#| label: reg_all_cause_mortality

#Create the regression model
reg_model_DIED <- svyglm(
  DIED ~ IBD + UlcerativeColitis + CrohnsDisease + AGE + FEMALE + RACE + ZIPINC_QRTL + Insurance + PL_NCHS + HOSP_REGION + HOSP_BEDSIZE + HOSP_LOCTEACH + elixsum + htn + CKD +
  obesity + dm + CPD + Dementia + LiverDis + PVD + Hyperlipidemia + CHF + Cancer + ObstructiveSleepApnea + ValvHD + Smoking + PriorStroke + PreviousPCI + PreviousCABG,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_DIED, exponentiate = T, label = reg_var_labels)

```

### Cardiogenic Shock:

```{r}
#| label: reg_CardioShock

#Create the regression model
reg_model_CardioShock <- svyglm(
  CardioShock ~ IBD + UlcerativeColitis + CrohnsDisease + AGE + FEMALE + RACE + ZIPINC_QRTL + Insurance + PL_NCHS + HOSP_REGION + HOSP_BEDSIZE + HOSP_LOCTEACH + elixsum + htn + CKD +
  obesity + dm + CPD + Dementia + LiverDis + PVD + Hyperlipidemia + CHF + Cancer + ObstructiveSleepApnea + ValvHD + Smoking + PriorStroke + PreviousPCI + PreviousCABG,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_CardioShock, exponentiate = T, label = reg_var_labels)

```

### Cardiac Arrest:

```{r}
#| label: reg_CardioArrest

#Create the regression model
reg_model_CardiacArrest <- svyglm(
  CardiacArrest ~ IBD + UlcerativeColitis + CrohnsDisease + AGE + FEMALE + RACE + ZIPINC_QRTL + Insurance + PL_NCHS + HOSP_REGION + HOSP_BEDSIZE + HOSP_LOCTEACH + elixsum + htn + CKD +
  obesity + dm + CPD + Dementia + LiverDis + PVD + Hyperlipidemia + CHF + Cancer + ObstructiveSleepApnea + ValvHD + Smoking + PriorStroke + PreviousPCI + PreviousCABG,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_CardiacArrest, exponentiate = T, label = reg_var_labels)

```

### Stroke:

```{r}
#| label: reg_Stroke

#Create the regression model
reg_model_Stroke <- svyglm(
  Stroke ~ IBD + UlcerativeColitis + CrohnsDisease + AGE + FEMALE + RACE + ZIPINC_QRTL + Insurance + PL_NCHS + HOSP_REGION + HOSP_BEDSIZE + HOSP_LOCTEACH + elixsum + htn + CKD +
  obesity + dm + CPD + Dementia + LiverDis + PVD + Hyperlipidemia + CHF + Cancer + ObstructiveSleepApnea + ValvHD + Smoking + PriorStroke + PreviousPCI + PreviousCABG,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_Stroke, exponentiate = T, label = reg_var_labels)

```

### Utilization of percutaneous coronary intervention:

```{r}
#| label: reg_PCI

#Create the regression model
reg_model_PCI <- svyglm(
  PCIProcedure ~ IBD + UlcerativeColitis + CrohnsDisease + AGE + FEMALE + RACE + ZIPINC_QRTL + Insurance + PL_NCHS + HOSP_REGION + HOSP_BEDSIZE + HOSP_LOCTEACH + elixsum + htn + CKD +
  obesity + dm + CPD + Dementia + LiverDis + PVD + Hyperlipidemia + CHF + Cancer + ObstructiveSleepApnea + ValvHD + Smoking + PriorStroke + PreviousPCI + PreviousCABG,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_PCI, exponentiate = T, label = reg_var_labels)

```

### Utilization of coronary artery bypass grafting:

```{r}
#| label: reg_CABG

#Create the regression model
reg_model_CABG <- svyglm(
  CABGProcedure ~ IBD + UlcerativeColitis + CrohnsDisease + AGE + FEMALE + RACE + ZIPINC_QRTL + Insurance + PL_NCHS + HOSP_REGION + HOSP_BEDSIZE + HOSP_LOCTEACH + elixsum + htn + CKD +
  obesity + dm + CPD + Dementia + LiverDis + PVD + Hyperlipidemia + CHF + Cancer + ObstructiveSleepApnea + ValvHD + Smoking + PriorStroke + PreviousPCI + PreviousCABG,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_CABG, exponentiate = T, label = reg_var_labels)

```

## Multivariable Linear Regression:

### Length of Stay:

```{r}
#| label: reg_length_of_stay

# Create the regression model
reg_model_LOS <- svyglm(
  LOS ~ IBD + UlcerativeColitis + CrohnsDisease + AGE + FEMALE + RACE + ZIPINC_QRTL + Insurance + PL_NCHS + HOSP_REGION + HOSP_BEDSIZE + HOSP_LOCTEACH + elixsum + htn + CKD + obesity + dm + CPD + Dementia + LiverDis + PVD + Hyperlipidemia + CHF + Cancer + ObstructiveSleepApnea + ValvHD + Smoking + PriorStroke + PreviousPCI + PreviousCABG,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model_LOS, exponentiate = F, label = reg_var_labels)

```

### Inflation Adjusted Total Charge:

```{r}
#| label: reg_total_charge

# Create the regression model for Total Charge (Inflation Adjusted)
reg_model_TOTCHG <- svyglm(
  adj_TOTCHG ~ IBD + UlcerativeColitis + CrohnsDisease + AGE + FEMALE + RACE + ZIPINC_QRTL + Insurance + PL_NCHS + HOSP_REGION + HOSP_BEDSIZE + HOSP_LOCTEACH + elixsum + htn + CKD +
    obesity + dm + CPD + Dementia + LiverDis + PVD + Hyperlipidemia + CHF + Cancer + ObstructiveSleepApnea + ValvHD + Smoking + PriorStroke + PreviousPCI + PreviousCABG,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model_TOTCHG, exponentiate = F, label = reg_var_labels)

```
