---
title: "In-hospital outcomes of patients with liver cirrhosis undergoing colorectal cancer surgery: an analysis of the National Inpatient Sample"
subtitle: "Analysis for RCOP 2025Apr-NIS-7-Gastro"
author:
  - name: Eliza Aisha
    email: elizaaisha1@gmail.com
    affiliations: 
      - id: DMC
        name: Dow University of Health Sciences
        department: Department of Medicine
        
  - name: Mariam Shahabi
    affiliations: 
      - id: DMC
        name: Dow University of Health Sciences
        department: Department of Medicine   
---

```{r}
#| label: libraries_setup
#| include: false

# Load libraries
library(tidyverse)
library(arrow)
library(survey)
library(gtsummary)

# Load external dependencies
source("./scripts/labels.R")

```

```{r}
#| label: data_setup
#| include: false
#| cache: false

# Global options
TESTING <- F

#Import data
if (TESTING) {
  dta <- arrow::open_dataset("./nis_data/NIS_CRC_LiverDis_testing.parquet") |> collect()
} else {
  dta <- arrow::open_dataset("./nis_data/NIS_CRC_LiverDis.parquet/") |> collect()
}

# Set factor references
dta <- dta |>
  mutate(
    Insurance = fct_relevel(Insurance, "Private"),
    RACE = fct_relevel(RACE, "White"),
    DIED = fct_relevel(DIED, "No"),
    LOS = as.numeric(LOS),
    adj_TOTCHG = as.numeric(adj_TOTCHG),
    Sepsis = as.factor(Sepsis),
    WoundInfection = as.factor(WoundInfection),
    # Abcess = as.factor(Abcess),
    Pneumonia = as.factor(Pneumonia),
    # VTE = as.factor(VTE),
    GIHemorrhage = as.factor(GIHemorrhage),
    # AMI = as.factor(AMI),
    AcuteRespiratoryFailure = as.factor(AcuteRespiratoryFailure),
    # MechanicalVentilation = as.factor(MechanicalVentilation),
    # Stroke = as.factor(Stroke),
    AcuteKidneyInjury = as.factor(AcuteKidneyInjury)
  )

# Set survey options for lonely PSUs
options(survey.lonely.psu = "adjust")
options(survey.adjust.domain.lonely = TRUE)

# Define survey design
dsgn <- svydesign(
  id = ~ HOSPID,
  strata = ~ NIS_STRATUM,
  weights = ~ TRENDWT,
  data = dta,
  nest = TRUE
)

# Subset data for specific condition
dsgn <- subset(dsgn,  ColorectalCancer == "Yes" & Colectomy == "Yes" & AGE >= 18)

# Clean up memory
rm(dta)
gc()

```


## Preamble:

-   **Reference Papers:**

    -   [Huang et al. 2025](https://www.nature.com/articles/s41598-024-84758-6)
    -   [Lee et al. 2018](https://bmjopen.bmj.com/content/8/7/e020511.long)
    
-   **Study Objective**: Examine the impact of liver cirrhosis on outcomes among inpatient admissions with colorectal cancer undergoing colectomy.

-   **Data Source**: Cross-sectional analysis of the National Inpatient Sample (NIS) from 2018 to 2020.

-   **Cohort Definition:** Patients were included if they met the following criteria

    -   Age ≥18 years

    -   Primary diagnosis of colorectal cancer

    -   Undergoing colectomy

    -   Non-missing data for liver cirrhosis

-   **Primary Outcomes**:

    -   In-hospital all-cause mortality

    -   Total hospital length of stay (days)

    -   Total charge, inflation adjusted to 2020 (\$)
    

-   **Secondary Outcomes**:

    -   Sepsis

    -   Pneumonia

    -   Gastrointestinal Hemorrhage

    -   Acute Respiratory Failure

    -   Acute Kidney Injury
    

-   **Covariates for Adjustment:** All models adjusted for the following:

    -   Patient Demographics: Age, sex (male and female), Race (White, Black, Hispanic, Asian or Pacific Islander, Native American and Other), Household income(Quartile 1, Quartile 2, Quartile 3, Quartile 4), insurance (Medicare, Medicaid, private, other)
    
    -   Hospital Characterstics: Type of admission (non-elective, elective), bed size of hospital (small, medium, large), location and teaching status of hospital (rural, urban non-teaching, urban teaching), location of the hospital (northeast, midwest or north central, south, west), type of surgery(laparoscopic, open, robotic colectomy)

    -   Comorbidities: Hypertension, diabetes mellitus, obesity, hyperlipidemia, coronary artery disease, peripheral vascular disease, congestive heart failure, chronic kidney disease, chronic pulmonary disease, anemia, fluid and electrolyte disorders, HIV/AIDS, weight loss, lymphoma, alcohol abuse, drug abuse, depression, psychoses, rheumatoid disease, valvular disease, paralysis, neurological disorders, coagulopathy

-   **Statistical Methods:**

    -   **Univariable Analysis:**

        -   Categorical variables: Compared using Pearson’s chi-square test with Rao & Scott adjustment

        -   Continuous variables: Analyzed using the design-based Kruskal–Wallis test, a non-parametric alternative suitable for survey-weighted data

    -   **Multivariable Analysis:**

        -   Binary outcomes (e.g., mortality): Survey-weighted logistic regression using the `quasibinomial` family

        -   Continuous outcomes (e.g., LOS, hospital charges): Survey-weighted linear regression using the `gaussian` family

        -   Results reported as adjusted odds ratios (aORs) or beta coefficients with 95% confidence intervals

-   **Software:** All statistical analyses were performed using the R Statistical Language (Version 4.5.0; R Foundation for Statistical Computing, Vienna, Austria), incorporating survey-weighted procedures via the *survey* package to account for the complex sampling design of NIS.

## Baseline Table:

```{r}
#| label: baseline
#| cache: false

# Create baseline table
tbl_svybaseline <- dsgn %>%
  tbl_svysummary(
    by = Cirrhosis,
    include = baseline_var,
    label = baseline_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p()

# Print baseline table
tbl_svybaseline

```

## Outcomes Table:

```{r}
#| label: outcomes_table

# Create outcomes table
tbl_svyoutcomes <- dsgn %>%
  tbl_svysummary(
    by = Cirrhosis,
    include = outcome_var,
    label = outcome_var_labels,
    missing = "no"
  ) %>%
  add_p()

# Print outcomes table
tbl_svyoutcomes

```

## Multivariable Logistic Regression:

### All-Cause Mortality:

```{r}
#| label: reg_all_cause_mortality

#Create the regression model
reg_model_DIED <- svyglm(
  DIED ~ Cirrhosis + Colectomy_Type + AGE + FEMALE + RACE + Insurance + ZIPINC_QRTL + HOSP_REGION + HOSP_BEDSIZE + HOSP_LOCTEACH + ELECTIVE +
    HTN + CKD + DM + CPD + PVD + Hyperlipidemia + CHF + Obesity + Anemia + HIV + Hypothyroidism + WeightLoss + Lymphoma + AlcoholAbuse + DrugAbuse +
    Depression + FluidDisorder + Psychoses + RheumatoidDisease + ValvularDisease + Paralysis + NeurologicalDisorders + Coagulopathy,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_DIED, exponentiate = T, label = reg_var_labels)

```

### Sepsis:

```{r}
#| label: reg_Sepsis

#Create the regression model
reg_model_Sepsis <- svyglm(
  Sepsis ~ Cirrhosis + Colectomy_Type + AGE + FEMALE + RACE + Insurance + ZIPINC_QRTL + HOSP_REGION + HOSP_BEDSIZE + HOSP_LOCTEACH + ELECTIVE +
    HTN + CKD + DM + CPD + PVD + Hyperlipidemia + CHF + Obesity + Anemia + HIV + Hypothyroidism + WeightLoss + Lymphoma + AlcoholAbuse + DrugAbuse +
    Depression + FluidDisorder + Psychoses + RheumatoidDisease + ValvularDisease + Paralysis + NeurologicalDisorders + Coagulopathy,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_Sepsis, exponentiate = T, label = reg_var_labels)

```

### Pneumonia:

```{r}
#| label: reg_Pneumonia

#Create the regression model
reg_model_Pneumonia <- svyglm(
  Pneumonia ~ Cirrhosis + Colectomy_Type + AGE + FEMALE + RACE + Insurance + ZIPINC_QRTL + HOSP_REGION + HOSP_BEDSIZE + HOSP_LOCTEACH + ELECTIVE +
    HTN + CKD + DM + CPD + PVD + Hyperlipidemia + CHF + Obesity + Anemia + HIV + Hypothyroidism + WeightLoss + Lymphoma + AlcoholAbuse + DrugAbuse +
    Depression + FluidDisorder + Psychoses + RheumatoidDisease + ValvularDisease + Paralysis + NeurologicalDisorders + Coagulopathy,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_Pneumonia, exponentiate = T, label = reg_var_labels)

```

### Gastrointestinal Hemorrhage:

```{r}
#| label: reg_GIHemorrhage

#Create the regression model
reg_model_GIHemorrhage <- svyglm(
  GIHemorrhage ~ Cirrhosis + Colectomy_Type + AGE + FEMALE + RACE + Insurance + ZIPINC_QRTL + HOSP_REGION + HOSP_BEDSIZE + HOSP_LOCTEACH + ELECTIVE +
    HTN + CKD + DM + CPD + PVD + Hyperlipidemia + CHF + Obesity + Anemia + HIV + Hypothyroidism + WeightLoss + Lymphoma + AlcoholAbuse + DrugAbuse +
    Depression + FluidDisorder + Psychoses + RheumatoidDisease + ValvularDisease + Paralysis + NeurologicalDisorders + Coagulopathy,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_GIHemorrhage, exponentiate = T, label = reg_var_labels)

```

### Acute Respiratory Failure:

```{r}
#| label: reg_AcuteRespiratoryFailure

#Create the regression model
reg_model_AcuteRespiratoryFailure <- svyglm(
  AcuteRespiratoryFailure ~ Cirrhosis + Colectomy_Type + AGE + FEMALE + RACE + Insurance + ZIPINC_QRTL + HOSP_REGION + HOSP_BEDSIZE + HOSP_LOCTEACH + ELECTIVE +
    HTN + CKD + DM + CPD + PVD + Hyperlipidemia + CHF + Obesity + Anemia + HIV + Hypothyroidism + WeightLoss + Lymphoma + AlcoholAbuse + DrugAbuse +
    Depression + FluidDisorder + Psychoses + RheumatoidDisease + ValvularDisease + Paralysis + NeurologicalDisorders + Coagulopathy,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_AcuteRespiratoryFailure, exponentiate = T, label = reg_var_labels)


```

### Acute Kidney Injury:

```{r}
#| label: reg_AcuteKidneyInjury

#Create the regression model
reg_model_AcuteKidneyInjury <- svyglm(
  AcuteKidneyInjury ~ Cirrhosis + Colectomy_Type + AGE + FEMALE + RACE + Insurance + ZIPINC_QRTL + HOSP_REGION + HOSP_BEDSIZE + HOSP_LOCTEACH + ELECTIVE +
    HTN + CKD + DM + CPD + PVD + Hyperlipidemia + CHF + Obesity + Anemia + HIV + Hypothyroidism + WeightLoss + Lymphoma + AlcoholAbuse + DrugAbuse +
    Depression + FluidDisorder + Psychoses + RheumatoidDisease + ValvularDisease + Paralysis + NeurologicalDisorders + Coagulopathy,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_AcuteKidneyInjury, exponentiate = T, label = reg_var_labels)

```

## Multivariable Linear Regression:

### Length of Stay:

```{r}
#| label: reg_length_of_stay

# Create the regression model
reg_model_LOS <- svyglm(
  LOS ~ Cirrhosis + Colectomy_Type + AGE + FEMALE + RACE + Insurance + ZIPINC_QRTL + HOSP_REGION + HOSP_BEDSIZE + HOSP_LOCTEACH + ELECTIVE +
    HTN + CKD + DM + CPD + PVD + Hyperlipidemia + CHF + Obesity + Anemia + HIV + Hypothyroidism + WeightLoss + Lymphoma + AlcoholAbuse + DrugAbuse +
    Depression + FluidDisorder + Psychoses + RheumatoidDisease + ValvularDisease + Paralysis + NeurologicalDisorders + Coagulopathy,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model_LOS, exponentiate = F, label = reg_var_labels)

```

### Inflation Adjusted Total Charge:

```{r}
#| label: reg_total_charge

# Create the regression model for Total Charge (Inflation Adjusted)
reg_model_TOTCHG <- svyglm(
  adj_TOTCHG ~ Cirrhosis + Colectomy_Type + AGE + FEMALE + RACE + Insurance + ZIPINC_QRTL + HOSP_REGION + HOSP_BEDSIZE + HOSP_LOCTEACH + ELECTIVE +
    HTN + CKD + DM + CPD + PVD + Hyperlipidemia + CHF + Obesity + Anemia + HIV + Hypothyroidism + WeightLoss + Lymphoma + AlcoholAbuse + DrugAbuse +
    Depression + FluidDisorder + Psychoses + RheumatoidDisease + ValvularDisease + Paralysis + NeurologicalDisorders + Coagulopathy,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model_TOTCHG, exponentiate = F, label = reg_var_labels)

```
