---
title: "Impact of Atrial Fibrillation on In-Hospital Outcomes in Infective Endocarditis Patients"
subtitle: "Analysis for RCOP NIS Cardio8"
author:
  - name: Eliza Aisha
    email: elizaaisha1@gmail.com
    affiliations: 
      - id: DMC
        name: Dow University of Health Sciences
        department: Department of Medicine
---


```{r}
#| label: libraries_setup
#| include: false

# Load libraries
library(tidyverse)
library(arrow)
library(survey)
library(gtsummary)

# Load external dependencies
source("./scripts/labels.R")

```

```{r}
#| label: data_setup
#| include: false
#| cache: false

# Global options
TESTING <- F

#Import data
if (TESTING) {
  dta <- arrow::open_dataset("./nis_data/NIS_Afib_IE_testing.parquet") |> collect()
} else {
  dta <- arrow::open_dataset("./nis_data/NIS_Afib_IE.parquet/") |> collect()
}

# Set factor references
dta <- dta |>
  mutate(
    IE_Afib = fct_relevel(IE_Afib, "Infective Endocarditis without Atrial Fibrillation"),
    Insurance = fct_relevel(Insurance, "Private"),
    RACE = fct_relevel(RACE, "White"),
    DIED = as.factor(DIED),
    LOS = as.numeric(LOS),
    ARF = as.factor(ARF),
    AIS = as.factor(AIS),
    AHF = as.factor(AHF)
  )

# Set survey options for lonely PSUs
options(survey.lonely.psu = "adjust")
options(survey.adjust.domain.lonely = TRUE)

# Define survey design
dsgn <- svydesign(
  id = ~ HOSPID,
  strata = ~ NIS_STRATUM,
  weights = ~ TRENDWT,
  data = dta,
  nest = TRUE
)

# Subset data for specific condition
dsgn <- subset(dsgn, IE == "Yes")

# Clean up memory
rm(dta)
gc()


```

## Preamble:

-   **Reference Papers:**

    -   [Sidhu et al. 2022](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9207776/pdf/main.pdf)

-   **Study Objective**: Examine the impact of atrial fibrillation on in-hospital outcomes among inpatient admissions with a principal diagnosis of infective endocarditis.

-   **Data Source**: Cross-sectional analysis of the National Inpatient Sample (NIS) from 2018 to 2020.

-   **Patient Selection**: Included all inpatient admissions with a principal diagnosis of infective endocarditis

-   **Infective Endocarditis + Atrial Fibrillation Categories**:

    -   Infective Endocarditis with Atrial Fibrillation

    -   Infective Endocarditis without Atrial Fibrillation

-   **Primary Outcomes**:

    -   In-hospital all-cause mortality

    -   Total hospital length of stay (days)

    -   Total charge, inflation adjusted to 2020 (\$)
    
-   **Secondary Outcomes**:

    -   Acute Renal Failure

    -   Acute Ischemic Stroke

    -   Acute Heart Failure
    
-   **Statistical Analysis**: Multiple logistic and linear regression to determine the independent association of atrial fibrillation with in-hospital outcomes, respectively, adjusted for:

    -   Demographics: Age, sex, race, residential income, insurance, and hospital region,

    -   Medical History: Hypertension, diabetes mellitus, hyperlipidemia, coronary artery disease, peripheral vascular disease, heart failure, chronic kidney disease, chronic obstructive pulmonary disease, liver disease, obstructive sleep apnea, anemia, obesity, alochol use disorder, and tobacco use disorder.

    -   Comorbidities: Charlson comorbidity index.

-   **Software:** All analyses were performed using R Version 4.4.1 (R Foundation for Statistical Computing, Vienna, Austria)

## Baseline Table:

```{r}
#| label: baseline
#| cache: true

# Create baseline table
tbl_svybaseline <- dsgn %>%
  tbl_svysummary(
    by = IE_Afib,
    include = baseline_var,
    label = baseline_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p() %>%
  add_overall()

# Print baseline table
tbl_svybaseline

```

## Outcomes Table:

```{r}
#| label: outcomes_table

# Create outcomes table
tbl_svyoutcomes <- dsgn %>%
  tbl_svysummary(
    by = IE_Afib,
    include = outcome_var,
    label = outcome_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p() %>%
  add_overall()

# Print outcomes table
tbl_svyoutcomes

```

## Multivariable Logistic Regression:

### All-Cause Mortality:

```{r}
#| label: reg_all_cause_mortality

# Create the regression model for all cause mortality
reg_model_died <- svyglm(
  DIED ~ IE_Afib + AGE + FEMALE + RACE + charlindex + ZIPINC_QRTL + Insurance + Hosp_Census_Region + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + Liver_Dis + OSA + Anemia + Alcohol + Obesity + Tobb,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_died, exponentiate = T, label = reg_var_labels)

```

### Acute Renal Failure:

```{r}
#| label: ARF

# Create the regression model for ARF
reg_model_ARF <- svyglm(
  ARF ~ IE_Afib + AGE + FEMALE + RACE + charlindex + ZIPINC_QRTL + Insurance + Hosp_Census_Region + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + Liver_Dis + OSA + Anemia + Alcohol + Obesity + Tobb,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_ARF, exponentiate = T, label = reg_var_labels)

```

### Acute Ischemic Stroke:

```{r}
#| label: AIS

# Create the regression model for AIS
reg_model_AIS <- svyglm(
  AIS ~ IE_Afib + AGE + FEMALE + RACE + charlindex + ZIPINC_QRTL + Insurance + Hosp_Census_Region + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + Liver_Dis + OSA + Anemia + Alcohol + Obesity + Tobb,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_AIS, exponentiate = T, label = reg_var_labels)

```

### Acute Heart Failure:

```{r}
#| label: AHF

# Create the regression model for AHF
reg_model_AHF <- svyglm(
  AHF ~ IE_Afib + AGE + FEMALE + RACE + charlindex + ZIPINC_QRTL + Insurance + Hosp_Census_Region + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + Liver_Dis + OSA + Anemia + Alcohol + Obesity + Tobb,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_AHF, exponentiate = T, label = reg_var_labels)

```

## Multivariable Linear Regression:

### Length of Stay:

```{r}
#| label: reg_length_of_stay

# Create the regression model
reg_model_LOS <- svyglm(
  LOS ~ IE_Afib + AGE + FEMALE + RACE + charlindex + ZIPINC_QRTL + Insurance + Hosp_Census_Region + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + Liver_Dis + OSA + Anemia + Alcohol + Obesity + Tobb,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model_LOS, exponentiate = F, label = reg_var_labels)

```

### Inflation Adjusted Total Charge:

```{r}
#| label: reg_total_charge

# Create the regression model
reg_model_TOTCHG <- svyglm(
  adj_TOTCHG ~ IE_Afib + AGE + FEMALE + RACE + charlindex + ZIPINC_QRTL + Insurance + Hosp_Census_Region + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + Liver_Dis + OSA + Anemia + Alcohol + Obesity + Tobb,
  design = dsgn,
  family = "gaussian"
  )

# Print regression table
tbl_regression(reg_model_TOTCHG, exponentiate = F, label = reg_var_labels)

```
