---
title: "Clinical and Economic Outcomes Associated with Aspiration Pneumonia in Hospitalized Patients with Parkinson’s Disease: A National Inpatient Sample Analysis"
subtitle: "Analysis for 2025April-NIS-16-IM"
author:
  - name: Eliza Aisha
    email: elizaaisha1@gmail.com
    affiliations: 
      - id: DMC
        name: Dow University of Health Sciences
        department: Department of Medicine    
---


```{r}
#| label: libraries_setup
#| include: false

# Load libraries
library(tidyverse)
library(arrow)
library(survey)
library(gtsummary)

# Load external dependencies
source("./scripts/labels.R")

```

```{r}
#| label: data_setup
#| include: false
#| cache: false

# Global options
TESTING <- F

 #Import data
 if (TESTING) {
   dta <- arrow::open_dataset("./data/NIS_PD_AP_testing.parquet") |> collect()
 } else {
   dta <- arrow::open_dataset("./data/NIS_PD_AP.parquet/") |> collect()
 }

# Set factor references
dta <- dta |>
  mutate(
    PD_AP = fct_relevel(PD_AP, "Parkinson's disease without aspiration pneumonia"),
    Insurance = fct_relevel(Insurance, "Private"),
    RACE = fct_relevel(RACE, "White"),
    DIED = as.factor(DIED),
    MVent = as.factor(MVent),
    Sepsis = as.factor(Sepsis),
    LOS = as.numeric(LOS),
    Disp_home = as.factor(Disp_home),
    Disp_short = as.factor(Disp_short),
    Disp_long = as.factor(Disp_long),
    Pall_Care = as.factor(Pall_Care),
    ARF = as.factor(ARF)
  )

# Set survey options for lonely PSUs
options(survey.lonely.psu = "adjust")
options(survey.adjust.domain.lonely = TRUE)

# Define survey design
dsgn <- svydesign(
  id = ~ HOSPID,
  strata = ~ NIS_STRATUM,
  weights = ~ TRENDWT,
  data = dta,
  nest = TRUE
)

# Subset data for specific condition
dsgn <- subset(dsgn, PD == "Yes" & !is.na(PD_AP))

# Clean up memory
rm(dta)
gc()


```

## Preamble:

-   **Reference Papers:**

    -   [Luca et al. 2021](https://journals.sagepub.com/doi/full/10.3233/JPD-212798)

-   **Study Objective**: Examine the impact of aspiration pneumonia on in-hospital outcomes among inpatient admissions with Parkinson's disease.

-   **Data Source**: Cross-sectional analysis of the National Inpatient Sample (NIS) from 2018 to 2020.

-   **Patient Selection**: Included all inpatient admissions with a diagnosis of Parkinson's disease.

-   **Aspiration Pneumonia + Parkinson's Disease Categories**:

    -   Parkinson's Disease with Aspiration Pneumonia

    -   Parkinson's Disease without Aspiration Pneumonia

-   **Outcomes**:

    -   In-hospital all-cause mortality

    -   Total hospital length of stay (days)

    -   Total charge, inflation adjusted to 2020 (\$)
    
    -   Sepsis

    -   Acute respiratory failure

    -   Mechanical Ventilation 
    
    -   Discharge Disposition (Discharged to home, short term facility, or long-term facility)
  
    -  Palliative care utilization

-   **Covariates for Adjustment:** All models adjusted for the following

    -   Demographics: Age, sex, race, residential income, insurance, hospital region, hospital bedsize, residence (urbanicity), hospital location, and teaching status.

    -   Medical History: Hypertension, diabetes mellitus, cancer, hyperlipidemia, coronary artery disease, peripheral vascular disease, heart failure, chronic kidney disease, chronic obstructive pulmonary disease, dementia, liver disease, fluid/electrolyte imbalance.

    -   Comorbidities: Elixhauser comorbidity index.

-   **Statistical Methods:**

    -   **Univariable Analysis:**

        -   Categorical variables: Compared using Pearson’s chi-square test with Rao & Scott adjustment

        -   Continuous variables: Analyzed using the design-based Kruskal–Wallis test, a non-parametric alternative suitable for survey-weighted data

    -   **Multivariable Analysis:**

        -   Binary outcomes (e.g., mortality): Survey-weighted logistic regression using the `quasibinomial` family

        -   Continuous outcomes (e.g., LOS, hospital charges): Survey-weighted linear regression using the `gaussian` family

        -   Results reported as adjusted odds ratios (aORs) or beta coefficients with 95% confidence intervals

-   **Software:** All statistical analyses were performed using the R Statistical Language (Version 4.5.0; R Foundation for Statistical Computing, Vienna, Austria), incorporating survey-weighted procedures via the *survey* package to account for the complex sampling design of NIS.

## Baseline Table:

```{r}
#| label: baseline
#| cache: true

# Create baseline table
tbl_svybaseline <- dsgn %>%
  tbl_svysummary(
    by = PD_AP,
    include = baseline_var,
    label = baseline_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p() %>%
  add_overall()

# Print baseline table
tbl_svybaseline

```

## Outcomes Table:

```{r}
#| label: outcomes_table

# Create outcomes table
tbl_svyoutcomes <- dsgn %>%
  tbl_svysummary(
    by = PD_AP,
    include = outcome_var,
    label = outcome_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p() %>%
  add_overall()

# Print outcomes table
tbl_svyoutcomes

```

## Multivariable Logistic Regression:

### All-Cause Mortality:

```{r}
#| label: reg_all_cause_mortality

# Create the regression model for all cause mortality
reg_model_died <- svyglm(
  DIED ~ PD_AP + AGE + FEMALE + RACE + elixsum + ZIPINC_QRTL + Insurance + Hosp_Census_Region + PL_NCHS +
    HTN + DM + Cancer + CHF + Hyperlip + CAD + PVD + CKD + COPD + Cardiac_arrhythmia + Dementia + Liver_Dis + Fluid_disorders,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_died, exponentiate = T, label = reg_var_labels)

```

### Acute Respiratory Failure:

```{r}
#| label: ARF

reg_model_ARF <- svyglm(
  ARF ~ PD_AP + AGE + FEMALE + RACE + elixsum + ZIPINC_QRTL + Insurance + Hosp_Census_Region + PL_NCHS +
    HTN + DM + Cancer + CHF + Hyperlip + CAD + PVD + CKD + COPD + Cardiac_arrhythmia + Dementia + Liver_Dis + Fluid_disorders,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_ARF, exponentiate = T, label = reg_var_labels)

```

### Sepsis:

```{r}
#| label: sepsis

reg_model_sepsis <- svyglm(
  Sepsis ~ PD_AP + AGE + FEMALE + RACE + elixsum + ZIPINC_QRTL + Insurance + Hosp_Census_Region + PL_NCHS +
    HTN + DM + Cancer + CHF + Hyperlip + CAD + PVD + CKD + COPD + Cardiac_arrhythmia + Dementia + Liver_Dis + Fluid_disorders,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_sepsis, exponentiate = T, label = reg_var_labels)

```

### Utilization of mechanical ventilation:

```{r}
#| label: MVent

reg_model_MVent <- svyglm(
  MVent ~ PD_AP + AGE + FEMALE + RACE + elixsum + ZIPINC_QRTL + Insurance + Hosp_Census_Region + PL_NCHS +
    HTN + DM + Cancer + CHF + Hyperlip + CAD + PVD + CKD + COPD + Cardiac_arrhythmia + Dementia + Liver_Dis + Fluid_disorders,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_MVent, exponentiate = T, label = reg_var_labels)

```

### Discharged to home:

```{r}
#| label: Dis_home

# Create the regression model for Disp_home
reg_model_Disp_home <- svyglm(
  Disp_home ~ PD_AP + AGE + FEMALE + RACE + elixsum + ZIPINC_QRTL + Insurance + Hosp_Census_Region + PL_NCHS +
    HTN + DM + Cancer + CHF + Hyperlip + CAD + PVD + CKD + COPD + Cardiac_arrhythmia + Dementia + Liver_Dis + Fluid_disorders,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_Disp_home, exponentiate = T, label = reg_var_labels)

```

### Transferred to short-term facility:

```{r}
#| label: Dis_short

reg_model_Disp_short <- svyglm(
  Disp_short ~ PD_AP + AGE + FEMALE + RACE + elixsum + ZIPINC_QRTL + Insurance + Hosp_Census_Region + PL_NCHS +
    HTN + DM + Cancer + CHF + Hyperlip + CAD + PVD + CKD + COPD + Cardiac_arrhythmia + Dementia + Liver_Dis + Fluid_disorders,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_Disp_short, exponentiate = T, label = reg_var_labels)

```

### Transferred to skilled nursing facility, intermediate care, or another facility:

```{r}
#| label: Disp_long

# Create the regression model for Disp_long
reg_model_Disp_long <- svyglm(
  Disp_long ~ PD_AP + AGE + FEMALE + RACE + elixsum + ZIPINC_QRTL + Insurance + Hosp_Census_Region + PL_NCHS +
    HTN + DM + Cancer + CHF + Hyperlip + CAD + PVD + CKD + COPD + Cardiac_arrhythmia + Dementia + Liver_Dis + Fluid_disorders,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_Disp_long, exponentiate = T, label = reg_var_labels)

```

### Utilization of Palliative Care:

```{r}
#| label: Pall_care

# Create the regression model for palliative care consultation
reg_model_Pall_Care <- svyglm(
  Pall_Care ~ PD_AP + AGE + FEMALE + RACE + elixsum + ZIPINC_QRTL + Insurance + Hosp_Census_Region + PL_NCHS +
    HTN + DM + Cancer + CHF + Hyperlip + CAD + PVD + CKD + COPD + Cardiac_arrhythmia + Dementia + Liver_Dis + Fluid_disorders,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_Pall_Care, exponentiate = T, label = reg_var_labels)

```

## Multivariable Linear Regression:

### Length of Stay:

```{r}
#| label: reg_length_of_stay

# Create the regression model
reg_model_LOS <- svyglm(
  LOS ~ PD_AP + AGE + FEMALE + RACE + elixsum + ZIPINC_QRTL + Insurance + Hosp_Census_Region + PL_NCHS +
    HTN + DM + Cancer + CHF + Hyperlip + CAD + PVD + CKD + COPD + Cardiac_arrhythmia + Dementia + Liver_Dis + Fluid_disorders,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model_LOS, exponentiate = F, label = reg_var_labels)

```

### Inflation Adjusted Total Charge:

```{r}
#| label: reg_total_charge

# Create the regression model
reg_model_TOTCHG <- svyglm(
  adj_TOTCHG ~ PD_AP + AGE + FEMALE + RACE + elixsum + ZIPINC_QRTL + Insurance + Hosp_Census_Region + PL_NCHS + HTN + DM + Cancer + CHF + Hyperlip + CAD + PVD + CKD + COPD + Cardiac_arrhythmia + Dementia + Liver_Dis + Fluid_disorders,
  design = dsgn,
  family = "gaussian"
  )

# Print regression table
tbl_regression(reg_model_TOTCHG, exponentiate = F, label = reg_var_labels)

```
