---
title: "Urban-Rural Disparities in Stroke Management and Outcomes Among Patients with Atrial Fibrillation"
subtitle: "Analysis for RCOP group NIS-Sep-Cardio5"
author:
  - name: Eliza Aisha
    email: elizaaisha1@gmail.com
    orcid: 0009-0007-7592-176X
    affiliations: 
      - id: DMC
        name: Dow University of Health Sciences
        department: Department of Medicine
---


```{r}
#| label: libraries_setup
#| include: false

# Load libraries
library(rmarkdown)
library(knitr)
library(yaml)
library(tidyverse)
library(arrow)
library(survey)
library(gtsummary)

# Load external dependencies
source("./scripts/labels.R")

```

```{r}
#| label: data_setup
#| include: false
#| cache: false

 # Global options
 TESTING <- F

 #Import data
 if (TESTING) {
   dta <- arrow::open_dataset("./nis_data/NIS_Stroke_Afib_Urban_testing.parquet") |> collect()
 } else {
   dta <- arrow::open_dataset("./nis_data/NIS_Stroke_Afib_Urban.parquet/") |> collect()
 }

 # Set factor references
dta <- dta |>
  mutate(
    Insurance = fct_relevel(Insurance, "Private"),
    RACE = fct_relevel(RACE, "White"),
    DIED = as.factor(DIED),
    LOS = as.numeric(LOS),
    ICH = as.factor(ICH),
    PL_NCHS = fct_relevel(PL_NCHS, "Rural"),
    MT = as.factor(MT)
  )

# Set survey options for lonely PSUs
options(survey.lonely.psu = "adjust")
options(survey.adjust.domain.lonely = TRUE)

# Define survey design
dsgn <- svydesign(
  id = ~ HOSPID,
  strata = ~ NIS_STRATUM,
  weights = ~ TRENDWT,
  data = dta,
  nest = TRUE
)

# Subset data for specific condition
dsgn <- subset(dsgn, Stroke == "Yes" & A_Fib == "Yes" & !is.na(PL_NCHS))

# Clean up memory
rm(dta)
gc()


```

## Preamble:

-   **Study Objective**: Examine impact of urbanization (urban vs. rural) on stroke management and outcomes among inpatient admissions with atrial fibrillation.

-   **Data Source**: Cross-sectional analysis of the National Inpatient Sample (NIS) from 2018 to 2020.

-   **Patient Selection**: Included all inpatient admissions with a principal diagnosis of Stroke.

-   **Outcomes**:

    -   In-hospital all-cause mortality

    -   Utilization of mechanical thrombectomy

    -   Intracranial hemorrhage

    -   Total hospital length of stay (days)

    -   Total charge, inflation adjusted to 2020 (\$)

-   **Statistical Analysis**: Multiple logistic and linear regression to determine the independent association of place of residence with in-hospital mortality, utilization of mechanical thrombectomy, intracranial hemorrhage, length of stay and total charges (inflation adjusted), respectively, adjusted for:

    -   Demographics: Age, sex, race, residential income, insurance, hospital region, hospital bedsize, hospital location, and teaching status.

    -   Medical History: Hypertension, diabetes mellitus, hyperlipidemia, coronary artery disease, peripheral vascular disease, heart failure, chronic kidney disease, chronic obstructive pulmonary disease, dementia, liver disease, fluid/electrolyte imbalance, obesity, prior stroke, prior cardiac surgery, presence of pacemaker/defibrillator.

    -   Comorbidities: Charlson comorbidity index.

-   **Software:** All analyses were performed using R Version 4.4.1 (R Foundation for Statistical Computing, Vienna, Austria)

## Baseline Table:

```{r}
#| label: baseline

# Create baseline table
tbl_svybaseline <- dsgn %>%
  tbl_svysummary(
    by = PL_NCHS,
    include = baseline_var,
    label = baseline_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p() %>%
  add_overall()

# Print baseline table
tbl_svybaseline

```

## Outcomes Table:

```{r}
#| label: outcomes_table

# Create outcomes table
tbl_svyoutcomes <- dsgn %>%
  tbl_svysummary(
    by = PL_NCHS,
    include = outcome_var,
    label = outcome_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p()

# Print outcomes table
tbl_svyoutcomes

```

## Multivariable Logistic Regression:

### All-Cause Mortality:

```{r}
#| label: reg_all_cause_mortality

# Create the regression model
reg_model_died <- svyglm(
  DIED ~ PL_NCHS + ZIPINC_QRTL + Insurance + AGE + FEMALE + RACE + charlindex + Insurance + Hosp_Census_Region + HOSP_BEDSIZE + HOSP_LOCTEACH + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + Dementia + Liver_Dis + Obesity + pStroke + pCardiacSurg + Pacemaker_defib + Fluid_disorders,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_died, exponentiate = T, label = reg_var_labels)

``` 

### Intracranial Hemorrhage:

```{r}
#| label: reg_ICH

reg_model_ICH <- svyglm(
  ICH ~ PL_NCHS + ZIPINC_QRTL + Insurance + AGE + FEMALE + RACE + charlindex + Insurance + Hosp_Census_Region + HOSP_BEDSIZE + HOSP_LOCTEACH + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + Dementia + Liver_Dis + Obesity + pStroke + pCardiacSurg + Pacemaker_defib + Fluid_disorders,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_ICH, exponentiate = T, label = reg_var_labels)

```

### Utilization of Mechanical Thrombectomy:

```{r}
#| label: reg_MT

reg_model_MT <- svyglm(
  MT ~ PL_NCHS + ZIPINC_QRTL + Insurance + AGE + FEMALE + RACE + charlindex + Insurance + Hosp_Census_Region + HOSP_BEDSIZE + HOSP_LOCTEACH + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + Dementia + Liver_Dis + Obesity + pStroke + pCardiacSurg + Pacemaker_defib + Fluid_disorders,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_MT, exponentiate = T, label = reg_var_labels)
```

## Multivariable Linear Regression:

### Length of Stay:

```{r}
#| label: reg_length_of_stay

# Create the regression model
reg_model_LOS <- svyglm(
  LOS ~ PL_NCHS + ZIPINC_QRTL + Insurance + AGE + FEMALE + RACE + charlindex + Insurance + Hosp_Census_Region + HOSP_BEDSIZE + HOSP_LOCTEACH + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + Dementia + Liver_Dis + Obesity + pStroke + pCardiacSurg + Pacemaker_defib + Fluid_disorders,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model_LOS, exponentiate = F, label = reg_var_labels)

```

### Inflation Adjusted Total Charge:

```{r}
#| label: reg_total_charge

# Create the regression model for Total Charge (Inflation Adjusted)
reg_model_TOTCHG <- svyglm(
  adj_TOTCHG ~ PL_NCHS + ZIPINC_QRTL + Insurance + AGE + FEMALE + RACE + charlindex + Insurance + Hosp_Census_Region + HOSP_BEDSIZE + HOSP_LOCTEACH + HTN + DM + Hyperlip + CAD + PVD + CHF +
    CKD + COPD + Dementia + Liver_Dis + Obesity + pStroke + pCardiacSurg + Pacemaker_defib + Fluid_disorders,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model_TOTCHG, exponentiate = F, label = reg_var_labels)

```
