---
title: "Outcomes in Heart Failure with Preserved vs. Reduced Ejection Fraction in Patients with Iron Deficiency Anemia (with propensity score matching)"
author:
  - name: Eliza Aisha
    email: elizaaisha1@gmail.com
    orcid: 0009-0007-7592-176X
    affiliations: 
      - id: DMC
        name: Dow University of Health Sciences
        department: Department of Medicine
---


```{r}
#| label: libraries_setup
#| include: false

library(tidyverse)    # dplyr, ggplot2, forcats, etc.
library(arrow)        # fast Parquet import
library(MatchIt)      # propensity-score matching
library(cobalt)       # balance diagnostics
library(survey)       # complex-survey analysis
library(gtsummary)    # publication-ready tables

# Load external dependencies
source("./scripts/labels.R")

# Global options
TESTING <- F

#Import data
if (TESTING) {
  dta <- arrow::open_dataset("./data/NIS_HF_IDA_testing.parquet") |> collect()
} else {
  dta <- arrow::open_dataset("./data/NIS_HF_IDA_2.parquet/") |> collect()
}

ps_covars <- c("AGE","FEMALE","RACE","ZIPINC_QRTL","Insurance","Hosp_Census_Region","elixsum","HTN","Hyperlip","CAD","PVD","COPD","Dementia","OSA","Alcohol","Valv_HD","Fluid_disorders","Pacemaker","Obesity","CA","Malnutrition","pCardiacSurg","NIS_STRATUM","TRENDWT")

dta <- dta %>%
  filter(!is.na(HF_Categories) & !is.na(AGE) & !is.na(FEMALE) & !is.na(RACE) & !is.na(ZIPINC_QRTL) & !is.na(Insurance)
         & !is.na(Hosp_Census_Region) & !is.na(HTN) & !is.na(Hyperlip) & !is.na(CAD)
         & !is.na(PVD) & !is.na(COPD) & !is.na(Dementia) & !is.na(OSA) & !is.na(Alcohol)
         & !is.na(Valv_HD) & !is.na(Dementia) & !is.na(CA) & !is.na(Fluid_disorders) & !is.na(Obesity) & !is.na(Pacemaker)
         & !is.na(pMI) & !is.na(Malnutrition) & !is.na(pCardiacSurg))

nis <- dta %>%
  mutate(
    # binary treatment indicator (factor with two levels)
    treat = factor(
      if_else(HF_Categories == "Heart failure with preserved ejection fraction", 1, 0),
      levels = c(0, 1),
      labels = c("Heart failure with reduced ejection fraction", "Heart failure with preserved ejection fraction")
    ),
    # reference levels & formats
    Insurance = fct_relevel(Insurance, "Private"),
    RACE      = fct_relevel(RACE, "White"),
    DIED      = factor(DIED)
  ) %>% drop_na(any_of(ps_covars))

# --------------------------------------------------------------------------- #
# 5.  Propensity-score matching (nearest‐neighbor, within strata)             #
# --------------------------------------------------------------------------- #
m.out <- matchit(
  treat ~ AGE + FEMALE + RACE + ZIPINC_QRTL + Insurance + Hosp_Census_Region + elixsum + HTN + Hyperlip + CAD + PVD + COPD + Dementia + OSA + Alcohol +
    Valv_HD + Fluid_disorders + Pacemaker + Obesity + CA + Obesity + Malnutrition + pCardiacSurg,
  data       = nis,
  method     = "nearest",
  distance   = "glm",
  exact      = ~ NIS_STRATUM,     # keep matches inside each stratum
  caliper    = 0.20,
  ratio      = 1,
  s.weights  = ~ TRENDWT          # survey weight enters PS model
)

matched <- match.data(m.out) %>%
  mutate(final_wt = weights * TRENDWT)  # `weights` is already present in `match.data(m.out)`

options(survey.lonely.psu = "adjust")           # HCUP recommendation
dsgn <- svydesign(
  id      = ~HOSPID,
  strata  = ~NIS_STRATUM,
  weights = ~final_wt,
  data    = matched,
  nest    = TRUE
)

```

## Preamble:

-   **Data Source**: Cross-sectional analysis of the National Inpatient Sample (NIS) from 2018 to 2020.

-   **Patient Selection**: Included all inpatient admissions with iron deficiency anemia.

-   **HF Categories**:

    -   Heart failure with reduced ejection fraction
    
    -   Heart failure with preserved ejection fraction

-   **ICD Codes**:
 
    -   HFrEF = I50.2 and I50.4
    
    -   HFpEF = I50.3
    
-   **Propensity score matching**:

    -   Using R package MatchIt
    
    -   Nearest neighbor matching, caliper = 0.2, ratio = 1:1    

-   **Outcomes**:

    -   In-hospital all-cause mortality

    -   Total hospital length of stay (days)

    -   Total charge, inflation adjusted to 2020 (\$)

-   **Software:** All analyses were performed using R Version 4.4.1 (R Foundation for Statistical Computing, Vienna, Austria)

## Baseline Table:

```{r}
#| label: baseline_table
#| cache: false

# Create baseline table
tbl_svybaseline <- dsgn %>%
  tbl_svysummary(
    by = HF_Categories,
    include = baseline_var,
    label = baseline_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p()

# Print baseline table
tbl_svybaseline


```

## Outcomes Table:

```{r}
#| label: outcomes_table

# Create outcomes table
tbl_svyoutcomes <- dsgn %>%
  tbl_svysummary(
    by = HF_Categories,
    include = outcome_var,
    label = outcome_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p()

# Print outcomes table
tbl_svyoutcomes

```


## Multivariable Logistic Regression:

### All-Cause Mortality:

```{r}
#| label: reg_all_cause_mortality

# Create the regression model for all cause mortality
reg_model_died <- svyglm(
  DIED ~ HF_Categories,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_died, exponentiate = T, label = reg_var_labels)
```

## Multivariable Linear Regression:

### Length of Stay:

```{r}
#| label: reg_length_of_stay

# Create the regression model
reg_model_LOS <- svyglm(
  LOS ~ HF_Categories,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model_LOS, exponentiate = F, label = reg_var_labels)

```

### Inflation Adjusted Total Charge:

```{r}
#| label: reg_total_charge

# Create the regression model for Total Charge (Inflation Adjusted)
reg_model_TOTCHG <- svyglm(
  adj_TOTCHG ~ HF_Categories,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model_TOTCHG, exponentiate = F, label = reg_var_labels)

```

