---
title: "Outcomes in HFrEF vs HFpEF in Patients with Iron Deficiency Anemia"
author:
  - name: Eliza Aisha
    email: elizaaisha1@gmail.com
    orcid: 0009-0007-7592-176X
    affiliations: 
      - id: DMC
        name: Dow University of Health Sciences
        department: Department of Medicine
---


```{r}
#| label: libraries_setup
#| include: false

# Load libraries
library(tidyverse)
library(arrow)
library(survey)
library(gtsummary)

# Load external dependencies
source("./scripts/labels.R")

# Global options
TESTING <- F

#Import data
if (TESTING) {
  dta <- arrow::open_dataset("./data/NIS_HF_IDA_testing.parquet") |> collect()
} else {
  dta <- arrow::open_dataset("./data/NIS_HF_IDA_2.parquet/") |> collect()
}

# Set factor references
dta <- dta |>
  mutate(
    HF_Categories = fct_relevel(HF_Categories, "Heart failure with preserved ejection fraction"),
    Insurance = fct_relevel(Insurance, "Private"),
    RACE = fct_relevel(RACE, "White"),
    DIED = as.factor(DIED),
    LOS = as.numeric(LOS)
  )

# Set survey options for lonely PSUs
options(survey.lonely.psu = "adjust")
options(survey.adjust.domain.lonely = TRUE)

# Define survey design
dsgn <- svydesign(
  id = ~ HOSPID,
  strata = ~ NIS_STRATUM,
  weights = ~ TRENDWT,
  data = dta,
  nest = TRUE
)


# Subset data for specific condition
dsgn <- subset(dsgn, IDA == "Yes" & !is.na(HF_Categories) & AGE > 18)

# Clean up memory
rm(dta)
gc()

```

## Preamble:

-   **Data Source**: Cross-sectional analysis of the National Inpatient Sample (NIS) from 2018 to 2020.

-   **Patient Selection**: Included all inpatient admissions with Iron deficiency anemia.

-   **HF Categories**:

    -   Heart failure with reduced ejection fraction
    
    -   Heart failure with preserved ejection fraction

-   **ICD Codes**:
 
    -   HFrEF = I50.2 and I50.4
    -   HFpEF = I50.3

-   **Primary Outcomes**:

    -   In-hospital all-cause mortality

    -   Total hospital length of stay (days)

    -   Total charge, inflation adjusted to 2020 (\$)

-   **Software:** All analyses were performed using R Version 4.4.1 (R Foundation for Statistical Computing, Vienna, Austria)

## Baseline Table:

```{r}
#| label: baseline_table
#| cache: false

# Create baseline table
tbl_svybaseline <- dsgn %>%
  tbl_svysummary(
    by = HF_Categories,
    include = baseline_var,
    label = baseline_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p()

# Print baseline table
tbl_svybaseline


```

## Outcomes Table:

```{r}
#| label: outcomes_table

# Create outcomes table
tbl_svyoutcomes <- dsgn %>%
  tbl_svysummary(
    by = HF_Categories,
    include = outcome_var,
    label = outcome_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p()

# Print outcomes table
tbl_svyoutcomes

```


### Multivariable Logistic Regression:

#### All-Cause Mortality:

```{r}
#| label: reg_all_cause_mortality

# Create the regression model for all cause mortality
reg_model_died <- svyglm(
  DIED ~ HF_Categories + AGE + FEMALE + RACE + elixsum  + ZIPINC_QRTL + Insurance + Hosp_Census_Region + HOSP_BEDSIZE + HOSP_LOCTEACH +
    HTN + Hyperlip + PVD + CAD + COPD + Valv_HD + Dementia + CA + Fluid_disorders + Afib + BMI + Pacemaker + pMI + Depression + pCardiacSurg,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_died, exponentiate = T, label = reg_var_labels)
```

### Multivariable Linear Regression:

#### Length of Stay:

```{r}
#| label: reg_length_of_stay

# Create the regression model
reg_model_LOS <- svyglm(
  LOS ~ HF_Categories + AGE + FEMALE + RACE + elixsum  + ZIPINC_QRTL + Insurance + Hosp_Census_Region + HOSP_BEDSIZE + HOSP_LOCTEACH +
    HTN + Hyperlip + PVD + CAD + COPD + Valv_HD + Dementia + CA + Fluid_disorders + Afib + BMI + Pacemaker + pMI + Depression + pCardiacSurg,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model_LOS, exponentiate = F, label = reg_var_labels)

```

#### Inflation Adjusted Total Charge:

```{r}
#| label: reg_total_charge

# Create the regression model for Total Charge (Inflation Adjusted)
reg_model_TOTCHG <- svyglm(
  adj_TOTCHG ~ HF_Categories + AGE + FEMALE + RACE + elixsum  + ZIPINC_QRTL + Insurance + Hosp_Census_Region + HOSP_BEDSIZE + HOSP_LOCTEACH +
    HTN + Hyperlip + PVD + CAD + COPD + Valv_HD + Dementia + CA + Fluid_disorders + Afib + BMI + Pacemaker + pMI + Depression + pCardiacSurg,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model_TOTCHG, exponentiate = F, label = reg_var_labels)

```

