---
title: "Impact of Weekend Admission on In-Hospital Outcomes in Patients with Acute Myocardial Infarction Undergoing Percutaneous Coronary Intervention"
subtitle: "Analysis for RCOP 2025Apr-NIS-Cardio2"
author:
  - name: Eliza Aisha
    email: elizaaisha1@gmail.com
    affiliations: 
      - id: DMC
        name: Dow University of Health Sciences
        department: Department of Medicine
---

```{r}
#| label: libraries_setup
#| include: false

# Load libraries
library(tidyverse)
library(easystats)
library(arrow)
library(survey)
library(gtsummary)
library(cardx)
library(broom.helpers)

# Load external dependencies
source("./scripts/labels.R")

```

```{r}
#| label: data_setup
#| cache: true
#| include: false

# Global options
TESTING <- F

if (TESTING) {
  dta <- arrow::open_dataset(
    "./nis_data/NIS_Weekend_AMI_PCI_testing.parquet"
  ) |>
    collect()
} else {
  dta <- arrow::open_dataset(
    "./nis_data/NIS_Weekend_AMI_PCI.parquet/"
  ) |>
    collect()
}

# Set factor references
dta <- dta |>
  mutate(
    RACE = fct_relevel(RACE, "White"),
    DIED = fct_relevel(DIED, "No"),
    LOS = as.numeric(LOS),
    adj_TOTCHG = as.numeric(adj_TOTCHG),
    AWEEKEND = fct_relevel(AWEEKEND, "Monday-Friday"),
    Mechanical_Ventilation = fct_relevel(Mechanical_Ventilation, "No"),
    Favorable_Discharge = fct_relevel(Favorable_Discharge, "No"),
    CS = as.factor(CS),
    Cardaic_arrest = as.factor(Cardaic_arrest),
    AcuteKidneyInjury = as.factor(AcuteKidneyInjury),
    GIHemorrhage = as.factor(GIHemorrhage),
    Stroke = as.factor(Stroke),
    Stroke_all = as.factor(Stroke_all),
    Pulm_catheter = as.factor(Pulm_catheter),
    AMI_type = as.factor(AMI_type),
    Sepsis = fct_relevel(Sepsis, "No"),
    UTI = fct_relevel(UTI, "No"),
    CABG = as.factor(CABG),
    MCS = as.factor(MCS),
    CoronaryArteryDissection = as.factor(CoronaryArteryDissection),
    CoronaryArteryPerforation = as.factor(CoronaryArteryPerforation),
    PostproceduralComplications = as.factor(PostproceduralComplications),
    VentArrhythmia = as.factor(VentArrhythmia),
    BloodTransfusion = as.factor(BloodTransfusion),
    IABP = as.factor(IABP),
    Hemodialysis = as.factor(Hemodialysis),
    RecurrentAMI = as.factor(RecurrentAMI)
    )

# Set survey options for lonely PSUs
options(survey.lonely.psu = "adjust")
options(survey.adjust.domain.lonely = TRUE)

# Define survey design
dsgn <- svydesign(
  id = ~HOSPID,
  strata = ~NIS_STRATUM,
  weights = ~TRENDWT,
  data = dta,
  nest = TRUE
)

# Subset data for specific condition
dsgn <- subset(
  dsgn,
  AGE >= 18 &
    AMI == "Yes" &
    PCI == "Yes" &
    !is.na(AWEEKEND)
)

# Clean up memory
rm(dta)
gc()

```

## Preamble:

-   **Reference Studies:**

    -   [Liu et al., 2024](https://www.nature.com/articles/s41598-024-59571-w#Sec11)

-   **Study Objective:** To evaluate the association between weekend admission and in-hospital outcomes among adult patients hospitalized for acute myocardial infarction who underwent percutaneous coronary intervention.

-   **Data Source:** A retrospective cross-sectional study using the National Inpatient Sample (NIS), years 2018–2020.

-   **Cohort Definition:** Hospitalizations were included if they met the following criteria

    -   Age ≥ 18 years

    -   Principal diagnosis of acute myocardial infarction

    -   Underwent percutaneous coronary intervention

    -   Non-missing values for admission timing

-   **Exposure of Interest:** Weekend admission, categorizing admissions as Saturday–Sunday (weekend) vs. Monday–Friday (weekday).

-   **Outcomes of Interest:**

    -   **Primary Outcome:**

        -   In-hospital mortality

    -   **Secondary Outcomes:**

        -   Length of stay (LOS), in days

        -   Inflation-adjusted total hospital charges (in 2020 USD)

        -   Mechanical ventilation use

        -   Favorable discharge (discharge to home or home health; all other dispositions were classified as unfavorable)
        
        -   Cardiogenic Shock

        -   Cardaic arrest
        
        -   Acute kidney injury

        -   GI hemorrhage
        
        -   Stroke
        
        -   Pulmonary artery catheterization
        
        -   Sepsis
        
        -   Urinary tract infection
        
        -   Undergoing CABG
        
        -   Mechanical circulatory support
        
        -   Coronary artery dissection
        
        -   Coronary artery perforation
        
        -   Postprocedural complications
        
        -   Ventricular arrhythmia
        
        -   Blood transfusion
        
        -   Recurrent MI during the hospital stay
        
        -   Intra-aortic balloon pump
        
        -   Hemodialysis

-   **Covariates for Adjustment:** All models adjusted for the following

    -   **Demographic Factors:** Age, sex, race/ethnicity

    -   **Socioeconomic Factors:** Primary payer, ZIP code–based income quartile

    -   **Clinical Factors:** Elixhauser comorbidity index, thrombolytic drugs given, type of acute myocardial infarction (STEMI or NSTEMI)

    -   **Hospital-Level Factors:** Hospital region, bed size, and teaching status

-   **Statistical Methods:**

    -   **Univariable Analysis:**

        -   Categorical variables: Compared using Pearson’s chi-square test with Rao & Scott adjustment

        -   Continuous variables: Analyzed using the design-based Kruskal–Wallis test, a non-parametric alternative suitable for survey-weighted data

    -   **Multivariable Analysis:**

        -   Binary outcomes (e.g., mortality, mechanical ventilation, discharge): Survey-weighted logistic regression using the `quasibinomial` family

        -   Continuous outcomes (e.g., LOS, hospital charges): Survey-weighted linear regression using the `gaussian` family

        -   Results reported as adjusted odds ratios (aORs) or beta coefficients with 95% confidence intervals

-   **Software:** All statistical analyses were performed using the R Statistical Language (Version 4.5.0; R Foundation for Statistical Computing, Vienna, Austria), incorporating survey-weighted procedures via the *survey* package to account for the complex sampling design of NIS.

## Baseline table:

```{r}
#| label: baseline
#| cache: true

# Create baseline table
tbl_svybaseline <- dsgn %>%
  tbl_svysummary(
    by = AWEEKEND,
    include = baseline_var,
    label = baseline_var_labels,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_p() |>
  add_overall()

# Print baseline table
tbl_svybaseline

```

## Univariable analysis:

```{r}
#| label: outcomes_table

# Create outcomes table
tbl_svyoutcomes <- dsgn %>%
  tbl_svysummary(
    by = AWEEKEND,
    include = outcome_var,
    label = outcome_var_labels,
    missing = "no"
  ) %>%
  add_p() |>
  add_overall()

# Print outcomes table
tbl_svyoutcomes

```

## Multivariable logistic regression:

### In-hospital mortality:

```{r}
#| label: reg_DIED

# Mortality
#Create the regression model
reg_model <- svyglm(
  DIED ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model, exponentiate = T, label = reg_var_labels)

```

### Mechanical ventilation:

```{r}
#| label: reg_Mechanical_Ventilation

# Mechanical_Ventilation
#Create the regression model
reg_model <- svyglm(
  Mechanical_Ventilation ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model, exponentiate = T, label = reg_var_labels)

```

### Favorable discharge:

```{r}
#| label: reg_Favorable_Discharge

# Favorable_Discharge
#Create the regression model
reg_model <- svyglm(
  Favorable_Discharge ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model, exponentiate = T, label = reg_var_labels)

```

### Cardiogenic Shock:

```{r}
#| label: reg_CS

#Create the regression model
reg_model_CS <- svyglm(
  CS ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_CS, exponentiate = T, label = reg_var_labels)

```

### Cardiac Arrest:

```{r}
#| label: reg_Cardaic_arrest

#Create the regression model
reg_model_Cardaic_arrest <- svyglm(
  Cardaic_arrest ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_Cardaic_arrest, exponentiate = T, label = reg_var_labels)

```

### Acute Kidney Injury:

```{r}
#| label: reg_AcuteKidneyInjury

#Create the regression model
reg_model_AcuteKidneyInjury <- svyglm(
  AcuteKidneyInjury ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_AcuteKidneyInjury, exponentiate = T, label = reg_var_labels)

```

### GI Hemorrhage:

```{r}
#| label: reg_GIHemorrhage

#Create the regression model
reg_model_GIHemorrhage <- svyglm(
  GIHemorrhage ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_GIHemorrhage, exponentiate = T, label = reg_var_labels)

```

### Stroke:

```{r}
#| label: reg_Stroke

#Create the regression model
reg_model_Stroke <- svyglm(
  Stroke_all ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_Stroke, exponentiate = T, label = reg_var_labels)

```

### Pulmonary artery catheterization:

```{r}
#| label: reg_Pulm_catheter

#Create the regression model
reg_model_Pulm_catheter <- svyglm(
  Pulm_catheter ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_Pulm_catheter, exponentiate = T, label = reg_var_labels)

```

### Sepsis:

```{r}
#| label: reg_Sepsis

#Create the regression model
reg_model_Sepsis <- svyglm(
  Sepsis ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_Sepsis, exponentiate = T, label = reg_var_labels)

```

### Urinary tract infection:

```{r}
#| label: reg__UTI

#Create the regression model
reg_model_UTI <- svyglm(
  UTI ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_UTI, exponentiate = T, label = reg_var_labels)

```

### Undergoing CABG:

```{r}
#| label: reg_CABG

#Create the regression model
reg_model_CABG <- svyglm(
  CABG ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_CABG, exponentiate = T, label = reg_var_labels)

```

### Mechanical circulatory support:

```{r}
#| label: reg_MCS

#Create the regression model
reg_model_MCS <- svyglm(
  MCS ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_MCS, exponentiate = T, label = reg_var_labels)

```

### Coronary artery dissection:

```{r}
#| label: reg_CoronaryArteryDissection

#Create the regression model
reg_model_CoronaryArteryDissection <- svyglm(
  CoronaryArteryDissection ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_CoronaryArteryDissection, exponentiate = T, label = reg_var_labels)

```

### Coronary artery perforation:

```{r}
#| label: reg_CoronaryArteryPerforation

#Create the regression model
reg_model_CoronaryArteryPerforation <- svyglm(
  CoronaryArteryPerforation ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_CoronaryArteryPerforation, exponentiate = T, label = reg_var_labels)

```

### Postprocedural complications:

```{r}
#| label: reg_PostproceduralComplications

#Create the regression model
reg_model_PostproceduralComplications <- svyglm(
  PostproceduralComplications ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_PostproceduralComplications, exponentiate = T, label = reg_var_labels)

```

### Ventricular arrhythmia:

```{r}
#| label: reg_VentArrhythmia

#Create the regression model
reg_model_VentArrhythmia <- svyglm(
  VentArrhythmia ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_VentArrhythmia, exponentiate = T, label = reg_var_labels)

```

### Blood transfusion:

```{r}
#| label: reg_BloodTransfusion

#Create the regression model
reg_model_BloodTransfusion <- svyglm(
  BloodTransfusion ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_BloodTransfusion, exponentiate = T, label = reg_var_labels)

```

### Recurrent MI:

```{r}
#| label: reg_RecurrentAMI

#Create the regression model
reg_model_RecurrentAMI <- svyglm(
  RecurrentAMI ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_RecurrentAMI, exponentiate = T, label = reg_var_labels)

```

### Intra-aortic balloon pump:

```{r}
#| label: reg_IABP

#Create the regression model
reg_model_IABP <- svyglm(
  IABP ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_IABP, exponentiate = T, label = reg_var_labels)

```

### Hemodialysis:

```{r}
#| label: reg_Hemodialysis

#Create the regression model
reg_model_Hemodialysis <- svyglm(
  Hemodialysis ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "quasibinomial"
)

# Print regression table
tbl_regression(reg_model_Hemodialysis, exponentiate = T, label = reg_var_labels)

```

## Multivariable linear regression:

### Length of Stay:

```{r}
#| label: reg_LOS

# Length of Stay
# Create the regression model
reg_model <- svyglm(
  LOS ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model, exponentiate = F, label = reg_var_labels)

```

### Inflation-adjusted total charge:

```{r}
#| label: reg_adj_TOTCHG

# Total Charge (Inflation Adjusted)
# Create the regression model
reg_model <- svyglm(
  adj_TOTCHG ~
    AWEEKEND +
    AMI_type +
    AGE +
    FEMALE +
    RACE +
    Thrombolytic_Drugs +
    PAY1_alt +
    ZIPINC_QRTL +
    HOSP_REGION +
    HOSP_BEDSIZE +
    HOSP_LOCTEACH +
    elixsum,
  design = dsgn,
  family = "gaussian"
)

# Print regression table
tbl_regression(reg_model, exponentiate = F, label = reg_var_labels)

```
